The Inconsistency in G\"odel's Ontological Argument: \\ A Success Story for AI in Metaphysics

### 2016-01-31 Papers/InProgress/IJCAI/submission.tex
Improve Section 4

In the typical workflow during an attempt to prove a conjecture with a theorem prover, it is customary to check the consistency of the axioms first. For if the axioms are inconsistent, anything (including the conjecture) would be trivially derivable in classical logic (\emph{ex falso quodlibet}). Surprisingly, when this routine check was performed on G\"odel's axioms \cite{C40}, the \textsc{Leo-II} prover claimed that the axioms were inconsistent. Unfortunately, the refutation generated by \textsc{Leo-II} was barely human-readable. The text file was 153 lines\footnote{Long lines with an average of 184 characters per line.} long and used a machine-oriented proof calculus (higher-order resolution \cite{W47}) and syntax (TPTP THF \cite{J22}). Part of the file is displayed in Fig.~\ref{LEO-Proof}.

@@

This can easily be seen: Axioms A1a in
Fig.~\ref{InconsistencyIsabelleK} is implied by Axiom Ax2 and the
highlighted footnote remark in Fig.~\ref{GoedelScript}. Axioms A2 and
A5 in Fig.~\ref{InconsistencyIsabelleK} correspond to Ax4 and Ax3 in
Fig.~\ref{GoedelScript}. The definitions of essence and necessary
existence are easy to identify. Thus, we conclude that the verified
inconsistency from Fig.~\ref{InconsistencyIsabelleK} does apply to 
G\"odel's original manuscript.

\subsection{Inconsistency of G\"odel's Axioms in \SFiveU}
Isabelle/HOL's Sledgehammer tool, which orchestrates calls to
external provers such as \textsc{Leo-II}, still
fails to detect the inconsistency of G\"odel's axioms (in the standard
embedding), while a direct modeling of the problem in TPTP THF syntax
in combination with a direct call of \textsc{Leo-II} succeeded. In
other words, without independent experiments directly in TPTP THF the
inconsistency would thus not have been detected.

%\marginpar{ToDo: Are we really supposed to use \S instead of ``Sec.''?}
On the other hand (and further confirming the claims from \S\ref{sec:improvedembedding}), 
the reconstruction in Isabelle/HOL with the improved embedding for \SFiveU was more efficient: 
the inconsistency could be detected
by \textsc{Leo-II} also when called via
Sledgehammer. Moreover, the result could subsequently be verified with
Metis even without the Empty Essence Lemma; cf. Fig.~\ref{Inconsistency_S5U}. 
% Interestingly, Metis' argument is different from that of
% \S\ref{sec:arg1} and \S\ref{sec:arg2}, since axiom A1b is not used.

>>>>

In the typical workflow during an attempt to prove a conjecture with a theorem prover, it is customary to check the consistency of the axioms first. For if the axioms are inconsistent, anything (including the conjecture) would be trivially derivable in classical logic (\emph{ex falso quodlibet}). Surprisingly, when this routine check was performed on G\"odel's axioms \cite{C40}, the \textsc{Leo-II} prover claimed that the axioms were inconsistent. Unfortunately, the refutation generated by \textsc{Leo-II} was barely human-readable. The text file was 153 lines\footnote{Long lines with an average of 184 characters per line.} long and used machine-oriented calculus (higher-order resolution \cite{W47}) and syntax (TPTP THF \cite{J22}). Part of the file is displayed in Fig.~\ref{LEO-Proof}.

@@

This can easily be seen: Axiom A1a in
Fig.~\ref{InconsistencyIsabelleK} is implied by Axiom Ax2 and the
highlighted footnote remark in Fig.~\ref{GoedelScript}. Axioms A2 and
A5 in Fig.~\ref{InconsistencyIsabelleK} correspond to Ax4 and Ax3 in
Fig.~\ref{GoedelScript}. The definitions of essence and necessary
existence are easy to identify. Thus, we conclude that the verified
inconsistency from Fig.~\ref{InconsistencyIsabelleK} does apply to 
G\"odel's original manuscript.

\subsection{Inconsistency of G\"odel's Axioms in \SFiveU}

Isabelle/HOL's Sledgehammer tool, which orchestrates calls to
external provers such as \textsc{Leo-II}, still
fails to detect the inconsistency of G\"odel's axioms in the standard
embedding of \SFive, while a direct modeling of the problem in TPTP THF syntax
in combination with a direct call of \textsc{Leo-II} succeeded. In
other words, without independent experiments with no mediation through Sledgehammer, the
inconsistency would not have been detected.

%\marginpar{ToDo: Are we really supposed to use \S instead of ``Sec.''?}
On the other hand (and further confirming the claims from \S\ref{sec:improvedembedding}), 
the reconstruction in Isabelle/HOL with the improved embedding for \SFiveU was more efficient: 
the inconsistency could be detected
by \textsc{Leo-II} also when called via
Sledgehammer. Moreover, the result could subsequently be verified with
Metis even without the Empty Essence Lemma (cf. Fig.~\ref{Inconsistency_S5U}). 
% Interestingly, Metis' argument is different from that of
% \S\ref{sec:arg1} and \S\ref{sec:arg2}, since axiom A1b is not used.
